Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Flow

### 2022-07-30 paper/paper.tex
Updates from Overleaf

Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects. In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that contribute to the mean flow. Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes. GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows \cite{2016PhRvL.116u4501M,2017JFM...810..412T, 2022arXiv220505513M}.

@@

\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
\begin{itemize}
    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
    \item Compare GQL performance at various cutoffs
    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
\end{itemize}
\subsection{Spiral Turbulence}
Meseguer et al (2009 PRE): 
\begin{itemize}
    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
\end{itemize}

The spiral turbulence regime with $\mu < 0$ provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_1$ and $\Reyn_2$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.

@@

\begin{table}[!h]
\caption{Simulations}
\label{tab:simulations}
\begin{tabular}{rlllllllll}
\hline
Run & $\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
\hline
A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
\hline
D & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
E & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Subcritical SPT\\
\hline
F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
\end{tabular}
\vspace*{-4pt}
\end{table}%%%End of the table

@@

\funding{Who paid Morgan? Who paid me?}

>>>>

Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects. In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that contribute to the mean flow. Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes (see \cite{2022arXiv220505513M} for a recent review). GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.

@@

\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
\begin{itemize}
    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
    \item Compare GQL performance at various cutoffs
    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
\end{itemize}
\subsection{Spiral Turbulence}
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../figs/reo_rei_lsb.pdf}
    \caption{A summary of spiral turbulence runs in the $\Reyn_o-\Reyn_i$ plane showing linear stability boundary. Solid circles shows points where we have conducted GQL analysis; crosses represent DNS steps along the path to subcritical behaviour. }
    \label{fig:LSB}
\end{figure}
%Meseguer et al (2009 PRE): 
%\begin{itemize}
%    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
%    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
%\end{itemize}

The spiral turbulence regime provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.
Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
From these two seed runs, we decrease $\Reyn_i$ in steps shown in figure~\ref{fig:LSB} where each point represents a DNS or GQL simulation run for one viscous time. The solid circles in figure~\ref{fig:LSB} represent runs where we performed GQL analyses.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
    \label{fig:urms_tz_rei700}
\end{figure}

\begin{table}[!h]
\caption{Simulations}
\label{tab:simulations}
\begin{tabular}{rlllllllll}
\hline
Run & $\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
\hline
A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
\hline
D & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
E & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Subcritical SPT\\
\hline
F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
\end{tabular}
\vspace*{-4pt}
\end{table}%%%End of the table

@@

\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. }
